---
title: "r彙整"
author: "H24071281_劉馨蔓"
date: "2020/4/30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### 1.創建
```{r}
 x<-3
 a<-c(3,4,2,7,10,23,8,9,11) ##新增序列
 a
 a[c(5,8)]<-a[c(8,5)] ##點交換順序
 a
 a[2:7]<-c(a[7:2]) ##一段序列交換順序
 a
```

### 2.運算
```{r}
 5 + 3
 5 * 3
 5 / 3
 5 ^ 3; 5 ** 3 ##次方
 5 %% 3 ##餘數
 5 %/% 3 ##商
 2^0.5; sqrt(2) ## 開根號
 abs(-2) ##絕對值
 exp(10)
 log(2)
 log2(2)
 log(10, base = 10); log10(10) ##相等
 log(10, base = exp(1)); log(10) ##相等
 factorial(3) ##n!
```

### 3.製造序列
```{r}
 seq(1,5,1) ##創造一個1到5的序列，每個數的間隔為1
 seq(1, 10, length.out = 6)##1到10的序列中共6個等差的值
 rep(seq(1,5,1),rep(3,5)) ##對1到5的序列重複5次，但先對每個數字重複3次
 more.colors<-c("red","yellow","blue","green","magenta","cyan")
 color.vec<-more.colors[rep(seq(1,3),times=4)+rep(seq(0,3),each=3)]
 color.vec
 unique(color.vec) ##計算序列中總共有幾個不同的值
 length(color.vec) ##算序列中有幾個值
 table(color.vec) ##統計每個不同的值出現的次數
 sort(color.vec,decreasing=TRUE) ##對序列排序，decrease:數字或字母從大排到小，increase數字或字母從小排到大
 d<-c(3,4,2,7,10,23,8,9,11)
 sort(d)
 sort(d, decreasing=TRUE)#order get the position of the number not exactly the number
 inc.order<-order(d)
 inc.order
 dec.order<-order(d,decreasing=TRUE)
 dec.order#order:how to get the number not the position
 d[inc.order]
 d[dec.order]
 d[dec.order][3]
 d[order(d,decreasing=TRUE)[3]]
 d[inc.order[length(d)-2]]
 a<-c(T,T,F,T,F) ##T=TRUE,F=FALSE
 length(a==TRUE)
 sum(a) 
 b<-c(2,NA,5,10)
 ##用is來判斷
 is.na(b)#是否有NA
 which(is.na(b))#NA是第幾個
 which(is.na(b)==1)
 ind.T<-which(is.na(b))
 b[-ind.T]#false
 which(!(is.na(b)))
 which(is.na(b)==0)
 A<-c(3,4,NA,12,0.2,NA,-2)
 B<-c(3,5,"red",4,NA)
 c(is.numeric(A),is.character(A),is.numeric(B),is.character(B))#判斷是否為數字/字串
 c(sum(is.na(A)),sum(is.na(B)))
 c(which(is.na(A)),which(is.na(B)))
 c(which(A<1),which(B<1))
```

### 4.統計
```{r}
 score0<-c(81.2,89.6,64.2,91.3,77.4,84.5,NA,91.7,63.5,84.8,
            NA,87.8,87.9,80.9,74.8,64.3,78.7,91.3,76.9,74.9,87.6,88.4)
 a<-c(mean(score0),max(score0),min(score0),sum(score0))
 b<-c(mean(score0,na.rm=TRUE),max(score0,na.rm=TRUE),min(score0,na.rm=TRUE),sum(score0,na.rm=TRUE))
 a;b#透過 na.rm=TRUE可以不計算NA值
 x <- c(1:25)
 y<-c(mean(x),var(x),median(x),quantile(x,c(0.1,0.5,0.7)))
 y1<-summary(x)
 x <- c(1:25, NA)
 par(mfrow = c(1, 2))
 boxplot(x)
 hist(x)
 runif(x)##random in uniform distribution
 rnorm(100,mean=0,sd=1)
 rm(x, y) ##拿掉x和y
 objects() ##看存在的東西
```

### 5.近似值
```{r}
 round(125.78253)
 round(125.78253, digits = 3)
 round(125.78253, digits = -2)
 round(83524587, digits = -3);   round(83524587, digits = -5)
 signif(125.78253, digits = 2)
 signif(83524587, digits = 3); signif(83524587, digits = 5)
 floor(125.78253)
 ceiling(125.78253)
 trunc(125.78253)
```

### 6.矩陣
```{r}
 a <- 1:5; b <- 6:10;c <- 11:15; d <- 16:20
 M1 <- rbind(a, b, c, d)
 M2 <- cbind(a, b, c, d)
 dim(M1)##算矩陣的維度
 X <- 1:20
 matrix(X, nrow = 4, ncol = 5)##建立矩陣,上到下再左到右
 matrix(X, nrow = 4, ncol = 5, byrow = TRUE)##建立矩陣,左到右再上到下
 colnames(M1) <- c("c1", "c2", "c3", "c4", "c5")##對column命名
 M1
 M1[3, 1] ##[列,行]
 M1[9] #從上往下數再由左到右第9個值
```
 
### 7.迴圈
```{r}
 ##for迴圈
 ##for(index in vector){
 ##
 ##}
 v<-c(1:6)
 for(k in v){
  print(2*k+1)
 }
 ##don't know how long
 ##when change the number in vector, make sure that the vector is empty,
 ##or the wrong number will still store in the vector
 odd1<-c()
 for(k in v){
  odd1<-c(odd1,2*k+1)
  print(odd1)
 }
 odd1
 #know how long
 odd2<-rep(NA,6)
 for(k in v){
  odd2[k]<-2*k+1
  print(odd2)
 }
 odd2
 ##while迴圈
 ##while(condition){
 ##  
 ##}
 k<-3
 while(k<=11){
 print(k)
 k<-k+ 1
 }
```

### 8.if
```{r}
 #if(condition){
 #  commands when TRUE
 #}

 #*
 #if(condition){
 #  commands when TRUE
 #}else{
 #  commands when FALSE
 #}

 #*
 #if(condition){
 #  commands
 #}else if(condition){
 #  commands
 #}else{
 #  commands
 #}
 ##there is no condition after"else"
 ##if there will be a condition after "else" please use "else if"

 #ifelse(a==b,TRUE,FALSE)
 basket<-c("r","b","y","y","y","r","r","b")
 n.basket<-c(0,0,0)
 for(i in 1:length(basket)){
  if(basket[i]=="r"){
    n.basket[1]<-n.basket[1]+1
  }else if(basket[i]=="b"){
    n.basket[2]<-n.basket[2]+1
  }else if(basket[i]=="y"){
  n.basket[3]<-n.basket[3]+1
  }
 }
```

### 9.repeat & break
```{r}
##a
 set.seed(20191025)##設定隨機種子
 a<-sample(1:6,1)
 repeat{
  if(length(unique(a))<6){
    a<-c(a,sample(1:6,1))
   }else{
    break
   }
 }
 print(length(a))
##b
 set.seed(20191025)
 s<-c()
 for(i in 1:100){
  s<-c(s,runif(1))
  if(sum(s)>3){
    print(i)
    break
   }
 } 
 ##做法a和b結果相同
```

### 10.function
```{r}
 ##cat & paste
 m<-c("blue","green","red","orange")
 paste(m)
 cat(m)
 ##set a function
 perfect.number<-function(n){
  k<-1
  res<-c()
  while(k<n){
    fac<-c()
    for(j in 1:k){
      if((k%%j)==0) fac<-c(fac,j)
    }
    if(k==sum(fac[-length(fac)])){
      cat(paste(k,"=",sep=" "),
          paste(fac[-length(fac)],collapse="+"),"\n")
      res<-c(res,k)
    }
    k<-k+1
   }
   return(k)
 }
 perfect.number(1000)
 ##list
 L<-list(a=2,b=c(1,3,5,7),c=c("red","blue"))
 L$a ## $:接元素的名稱
 L[2]
 L[[2]] ##[[]]:接元素的位置或名稱
 L[["c"]]
```

### 11.dataframe
```{r}
 #as.data.frame()   short time
 x<-matrix(NA,3,3)
 x[,1]<-c("red","yellow","blue")
 x[,2]<-c(1,2,3)
 x[,3]<-c(4,5,6)
 colnames(x)<-c("colors","numbers","more.numbers")
 x
 y<-data.frame(x)
 y
 y$colors
 y$numbers
 #x$colors    matrix cannot use $ but dataframe can
 colors<-c("red","yellow","blue")
 numbers<-c(1,2,3)
 colors.and.numbers<-data.frame(colors,numbers,more.numbers=c(4,5,6),
                               stringsAsFactors = FALSE)
 str(colors.and.numbers)
 #data1<-data.frame(colors,numbers,more.numbers=c(4,5,6))
 #data1<-rbind(data1,c("black",3,5))     compare with data2
 data2<-data.frame(colors,numbers,more.numbers=c(4,5,6),
                  stringsAsFactors = FALSE)
 data2<-rbind(data2,c("black",3,5)) 
 data2
 str(data2)
```


### 12.畫圖
```{r, fig.width = 10, fig.height = 8}
par(mfrow=c(2,2))
x<- c(0: 10)
#1
plot(x, dbinom(x, size = 10, prob = .2), xlim = c(0, 10), ylim = c(0, 0.4), type = "h",
     ylab = "pmf", col = "red", lty = 8)
par(new = TRUE)
plot(x, dbinom(x, size = 10, prob = .2), xlim = c(0, 10), ylim = c(0, 0.4), type = "p",
      xaxt = "n", yaxt = "n", xlab = "", ylab = "", col = "red", pch = 17)
par(new = TRUE)
plot(x, dbinom(x, size = 10, prob = .5), xlim = c(0, 10), ylim = c(0, 0.4), type = "h",
      xaxt = "n", yaxt = "n", xlab = "", ylab = "", col = "blue", lty = 4)
par(new = TRUE)
plot(x, dbinom(x, size = 10, prob = .5), xlim = c(0, 10), ylim = c(0, 0.4), type = "p",
      xaxt = "n", yaxt = "n", xlab = "", ylab = "", col = "blue", pch = 8)
par(new = TRUE)
plot(x, dbinom(x, size = 10, prob = .7), xlim = c(0, 10), ylim = c(0, 0.4), type = "h",
      xaxt = "n", yaxt = "n", xlab = "", ylab = "", col = "black")
par(new = TRUE)
plot(x, dbinom(x, size = 10, prob = .7), xlim = c(0, 10), ylim = c(0, 0.4), type = "p",
      xaxt = "n", yaxt = "n", xlab = "", ylab = "",col="black", pch = 16)
legend( "topright", c("p=0.7", "p=0.5", "p=0.2"),
       col = c("black", "blue", "red"), lty = c( 1, 4, 8),pch = c(16, 8, 17))
title(expression("Binomial(n, p) with n = 10"))
#2
curve(dgamma(x, shape = 1, rate = 3, log = FALSE),xlim = c(0, 5), ylim = c(0, 3.0),
      ylab = "pdf")
par(new = TRUE)
curve(dgamma(x, shape = 2, rate = 3, log = FALSE),xlim = c(0, 5), ylim = c(0, 3.0),
      xaxt = "n", yaxt = "n", xlab = "", ylab = "", col = "red", lty = 2)
par(new = TRUE)
curve(dgamma(x, shape = 3, rate = 3, log = FALSE),xlim = c(0, 5), ylim = c(0, 3.0),
      xaxt = "n", yaxt = "n", xlab = "", ylab = "", col = "blue", lty = 4)
par(new = TRUE)
curve(dgamma(x, shape = 5, rate = 3, log = FALSE),xlim = c(0, 5), ylim = c(0, 3.0),
      xaxt = "n", yaxt = "n", xlab = "", ylab = "", col = "gray", lty = 11)
legend("topright", c(expression(paste(alpha, "=1")), expression(paste(alpha, "=2")),
                     expression(paste(alpha, "=3")),expression(paste(alpha, "=5"))),
       col = c("black", "red","blue","grey"), lty = c(1, 2, 4, 11))
title(expression(paste("Gamma(", alpha , ", ", lambda, ") with ", lambda, " = 3")))
#3
curve(dnorm(x, mean = 5, sd = 1), xlim = c(0, 20), ylim = c(0, 0.4), ylab = "pdf",
      lab = c(3, 5, 7))
par(new = TRUE)
curve(dnorm(x, mean = 10, sd = 1), xlim = c(0, 20), ylim = c(0, 0.4),
      xaxt = "n", yaxt = "n", xlab = "", ylab = "", col = "red", lty = 2)
par(new = TRUE)
curve(dnorm(x, mean = 12, sd = 1), xlim = c(0, 20), ylim = c(0, 0.4),
      xaxt = "n", yaxt = "n", xlab = "", ylab = "", col = "blue", lty = 4)
par(new = TRUE)
curve(dnorm(x, mean = 14, sd = 1), xlim = c(0, 20), ylim = c(0, 0.4),
      xaxt = "n", yaxt = "n", xlab = "", ylab = "", col = "grey", lty = 11)
legend("topright", c(expression(paste(mu, "=5")), expression(paste(mu, "=10")),
                     expression(paste(mu, "=12")),expression(paste(mu, "=14"))),
       col = c("black", "red", "blue", "grey"), lty = c(1, 2, 4, 11))
title(expression(paste("N(", mu, ", ", sigma^2, ") with ",sigma^2, " =1 ")))
#4
curve(dnorm(x, mean = 0, sd = 0.5), xlim = c(-8, 8), ylim = c(0, 0.8),ylab = "pdf",
      lab = c(3, 5, 7))
par(new = TRUE)
curve(dnorm(x, mean = 0, sd = 1), xlim = c(-8, 8), ylim = c(0, 0.8),
      xaxt = "n", yaxt = "n", xlab = "", ylab = "", col = "red", lty = 2)
par(new = TRUE)
curve(dnorm(x, mean = 0, sd = 2), xlim = c(-8, 8), ylim = c(0, 0.8),
      xaxt = "n", yaxt = "n", xlab = "", ylab = "", col = "blue", lty = 4)
par(new = TRUE)
curve(dnorm(x, mean = 0, sd = 3), xlim = c(-8, 8), ylim = c(0, 0.8),
      xaxt = "n", yaxt = "n", xlab = "", ylab = "", col = "grey", lty = 11)
legend("topright", c(expression(paste(sigma, "=0.5")), expression(paste(sigma, "=1")),
                     expression(paste(sigma, "=2")), expression(paste(sigma, "=3"))),
       col = c("black", "red","blue","grey"), lty = c(1, 2, 4, 11))
title(expression(paste("N(", mu, ", ", sigma^2, ") with ", mu, " = 0")))

``` 

### 13.rmarkdown

$$\bar{x} = \frac{\sum_{i=1}^{n} x_i}{n} $$

$$\alpha, \beta, \theta, \lambda, \pi, \tau, \Gamma, \gamma, \rho$$
Define $X = A+B$

* unordered list
* item 2
  + sub-item 1
  + sub 2
  
1. 123
2. 456
  a. 1
  b. 2

Table Header | Second Header
------------- | -------------
Table Cell | Cell 2
Cell 3 | Cell 4

\begin{table}[h]
    \centering
    \begin{tabular}{c|cc}
      \hline
      $x_j$ & 0 & 1 \\
      \hline
      $p_j$ & 0.3 & 0.7 \\
      \hline
    \end{tabular}
    \caption{Random variable $X$ with its probability}
\end{table}

$\left[
\begin{array}{cc}
A & B \\
C & D \\
\end{array}
\right]$


14 times 16 equals `r 14 * 16`.

### 14. evaluation of matrix
```{r}
A<-matrix(c(0:3),2,2)
A
A+5
A-2
B<-matrix(c(6:9),2,2)
B
A+B
A*B #element times element, eg:[1,1]*[1,1]
A %*% B #matix times matix, eg: matrix A *matix B
C<-matrix(c(1:9),3,3)
C
colSums(C);rowSums(C)
colMeans(C);rowMeans(C)
t(A) #transpose of A
diag(A) #diagnose ofA
det(A) #determinent of A
solve(A)
D <- matrix(rnorm(9), 3 ,3)
D
eig.D <- eigen(D)
eig.D$values
eig.D$vectors
```